---
title: "Assignment 4- Task 2 Mini Exploratory Report"
author: "Sydney Rilum"
date: "11/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(kableExtra)
library(lubridate)
```

## Introduction

A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last.


## Data and Analyses

A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.

```{r}
bonanza <- read.csv(here("data", "bonanza_hares.csv"))
```


## Exploratory Findings

### A. Annual juvenile hare trap counts


### B. Visualize juvenile hare weights


### C. Juvenile weight comparison (male & female snowshoe hares)


### D. Relationship between juvenile weight & hind foot length


## Summary

Include a very short summary describing your main findings from your juvenile hares exploration (bulletpoints are fine)


## Citations

Include citations (professionally formatted), including the data citation and any other resources that were references in your mini report. 


