---
title: "Juvenile Snowshoe Hares: Exploratory Report"
author: "Sydney Rilum"
date: "11/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(kableExtra)
library(lubridate)
library(ggbeeswarm)
library(effsize)
library(broom)
```

## Introduction

A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. 

This report provides an exploratory overview of juvenile snowshoe hare counts data from Bonanza Creek Experimental Forest, Alaska across three different sampling sites from 1998-2012 (Kielland et al. 2017). Specifically, juvenile hare weights are compared by sex and sampling site through descriptive statistics and data visualizations. In addition, the relationship between juvenile weight and hind foot length is explored.


## Data and Analyses

Snowshoe hare physical measurements were collected and made available by Kielland et al. at the Bonanza Creek Long Term Ecological Research Site in Alaska (Kielland et al. 2017). The data contains observations for 3,197 recorded hare trappings, collected across three sample sites ("Bonanza Black Spruce", "Bonanza Mature", and "Bonanza Riparian") differing in habitat and vegetation type, from the years 1998 to 2012. The following exploratory data visualization includes comparison of juvenile hare weights by sex and site through descriptive statistics, exploratory plots, a two-sample t-test using a significance level (α) of 0.05, and analysis of effect size by Cohen's d. The relationship between juvenile weight and hind foot length is explored by simple linear regression. All analyses are in R version 4.0.2 using RStudio version 1.3.1093


```{r}
# Read in the data
bonanza <- read.csv(here("data", "bonanza_hares.csv"))
```

## Exploratory Findings

### A. Annual juvenile hare trap counts

The total number of juvenile hare trappings during each year of the study were calculated in order to visualize the number of observations within the data set.

```{r}
# Create a subset of data that filters for juvenile hares and counts by year
ann_juvenile_counts <- bonanza %>% 
  mutate(date = mdy(date), .keep = "unused") %>% 
  mutate(year = year(date)) %>% 
  filter(age == "j") %>% 
  count(year)

# Calculate mean annual number of juvenile hares trapped
mean_juvenile_traps <- ann_juvenile_counts %>% 
  summarize(mean = mean(n))

# Create a bar graph displaying annual juvenile hare trap counts per year
ggplot(data = ann_juvenile_counts, aes(x = year, y = n)) +
  geom_col(fill = "cyan4") +
  labs(x = "Year", 
       y = "Number of juvenile hare trappings",
       title = "Annual juvenile hare trap counts") +
   scale_x_continuous(breaks = c(1999, 2001, 2003, 2005, 2007, 2009, 2011)) +
  theme_minimal() +
  theme(legend.position = "none")
```

**Figure 1.** Annual number of juvenile snowshoe hare trappings per year from 1999 to 2012. Note there are no juvenile hare traps recorded in 2002 and 2009. Data: Kielland et al. (2017).

As displayed in the bar graph, the maximum count of juvenile hare trappings was recorded in 1999 with 126 hare traps. The years 2004, 2007, 2010, and 2012 had the smallest number of juvenile hare trappings, with 8, 5, 2, and 6 hare traps respectively. In addition, the mean annual number of juvenile hare trappings was 31.5. Snowshoe hare densities appear to be highest in 1999 and  declining since. This is not atypical for snowshoe hares as they usually experience population fluctuations of 8-11 years (Kielland et al. 2017). It is important to note that these yearly counts can be impacted by effort. Thus, in order to standardize the juvenile hare population in future analyses, I would suggest that there is a consistent number of traps set for the same number of days each year in the study.  


### B. Visualize juvenile hare weights

Various data visualizations were created to compare juvenile hare weights by both sex and site, as well as by sex and site separately. When comparing juvenile hare weights by both site and sex in Figure 2, male hares have larger median weights than female hares across all three sampling sites. When hare weights are compared by sex alone in Figure 3, male hares are on average heavier than female hares, with means of 945.86 g and 855.39 g, respectively. Lastly, when sites alone are used to compare hare weights in Figure 4, the Bonanza Black Spruce site had the heaviest juvenile hares, with a mean weight of 1036.17 g. Juvenile hares at the Bonanza Mature and Bonanza Riparian sites had similar smaller mean weights of 878.05 g and 862.59 g, respectively.

```{r}
# Create a subset of data that filters for juvenile hares and renames sites
hare_weights <- bonanza %>% 
  filter(age == "j") %>% 
  select(weight, sex, grid) %>% 
  mutate(grid = case_when(
    grid == "bonbs" ~ "Bonanza Black Spruce", 
    grid == "bonmat" ~ "Bonanza Mature",
    grid == "bonrip" ~ "Bonanza Riparian"))

# Juvenile Hare Weights by Sex and Site
ggplot() +
  geom_beeswarm(data = hare_weights, 
                aes(x = sex, y = weight, color = grid),
                size = 2,
                alpha = 0.5) +
  geom_boxplot(data = hare_weights,
               aes(x = sex, y = weight),
               alpha = 0) +
  labs(x = "Sex", 
       y = "Weight (g)",
       title = "Juvenile Hare Weights by Sex and Site") +
  facet_wrap(~grid) +
  scale_color_manual(values = c("cornflowerblue","cyan4", "coral2")) +
  theme_light() +
  theme(legend.position = "none")
```

**Figure 2.** Juvenile snowshoe hare weights (g) compared by sex for each of the three sampling sites. A boxplot is overlaid on the beeswarm plot to display the distributions of each subset of data, including the median (bold vertical line) and 25^th^ and 75^th^ percentile values (box endpoints) for each subset. For the sex variable, "f" represents female hares, "m" represents male hares, and "NA" represents hare trap observations without a sex recorded. Data: Kielland et al. (2017).


```{r}
# Create a subset of data that calculates mean hare weight for each sex
hare_mean_weight_sex <- hare_weights %>% 
  group_by(sex) %>% 
  summarize(mean_weight = mean(weight, na.rm = TRUE))

# Juvenile Hare Weights by Sex
ggplot() +
  geom_beeswarm(data = hare_weights, 
                aes(x = sex, y = weight, color = sex),
                size = 2,
                alpha = 0.5) +
  geom_boxplot(data = hare_weights,
               aes(x = sex, y = weight),
               width = 0.3,
               alpha = 0) +
  geom_point(data = hare_mean_weight_sex, 
             aes(x = sex, y = mean_weight),
             size = 3,
             color = "black") +
  labs(x = "Sex", 
       y = "Weight (g)",
       title = "Juvenile Hare Weights by Sex") +
  theme_light() +
  theme(legend.position = "none")
```

**Figure 3.** Juvenile snowshoe hare weights (g) compared by sex. Box endpoints indicate the 25^th^ and 75^th^ percentile values; the black line and black point within the box indicate the median and mean value for each species, respectively. For the sex variable, "f" represents female hares, "m" represents male hares, and "NA" represents hare trap observations without a sex recorded. Data: Kielland et al. (2017).


```{r}
# Create a subset of data that calculates mean hare weight for each site
hare_mean_weight_site <- hare_weights %>% 
  group_by(grid) %>% 
  summarize(mean_weight = mean(weight, na.rm = TRUE))

# Juvenile Hare Weights by Site
ggplot() +
  geom_beeswarm(data = hare_weights, 
                aes(x = grid, y = weight, color = grid),
                size = 2,
                alpha = 0.5) +
  geom_boxplot(data = hare_weights,
               aes(x = grid, y = weight),
               width = 0.3,
               alpha = 0) +
  geom_point(data = hare_mean_weight_site, 
             aes(x = grid, y = mean_weight),
             size = 3,
             color = "black") +
  labs(x = "Site", 
       y = "Weight (g)",
       title = "Juvenile Hare Weights by Site") +
  theme_light() +
  theme(legend.position = "none")
```

**Figure 4.** Juvenile snowshoe hare weights (g) compared by sampling site. Box endpoints indicate the 25^th^ and 75^th^ percentile values; the black line and black point within the box indicate the median and mean value for each species, respectively. Data: Kielland et al. (2017).


### C. Juvenile weight comparison (male & female snowshoe hares)

```{r}
# Descriptive statistics data subset
hare_weight_stats <- bonanza %>% 
  filter(age == "j") %>% 
  mutate(sex = case_when(
    sex == "f" ~ "Female", 
    sex == "m" ~ "Male")) %>% 
  select(weight, sex) %>% 
  group_by(sex) %>% 
  summarize(mean_weight = round(mean(weight, na.rm = TRUE),2),
            sd_weight = round(sd(weight, na.rm = TRUE),2),
            sample_size = n())
```

**Table 1.** Descriptive statistics (mean, standard deviation, and sample size) for male and female juvenile snowshoe hares. Data: Kielland et al. (2017).

```{r}
# Descriptive statistics table
hare_weight_stats %>% 
  kable(col.names = c("Sex", "Mean weight (g)", "Standard deviation (g)", "Sample size")) %>% 
  kable_styling(bootstrap_options = "striped")
```

 
```{r, include = FALSE}
# T-test code: 

# Create vectors of weight for female and male juvenile hares separately.
hare_f <- hare_weights %>% 
  filter(sex == "f") %>% 
  pull(weight)

hare_m <- hare_weights %>% 
  filter(sex == "m") %>% 
  pull(weight)

# Plot histograms & qq-plots to check assumptions
hist(hare_f) # Looks normal
qqnorm(hare_f) # Looks very linear
hist(hare_m) # Possibly a bit negatively skewed
qqnorm(hare_m) # Looks relatively linear

# Overall: assumptions of normality hold, & sample size additionally >> Central Limit Theorem (means will be normally distributed). Two-sample t-test OK for means comparison! 


## T-test: Is there a significant difference in weight between male and female juvenile hares?

# Run the t-test:
hare_ttest <- t.test(hare_f, hare_m)
hare_ttest

# Get the tidy model results to call outputs in-line: 
hare_ttest_tidy <- tidy(hare_ttest)

# Get the means & SDs for each: 
male_mean <- mean(hare_m)
female_mean <- mean(hare_f, na.rm = TRUE)
male_sd <- sd(hare_m)
female_sd <- sd(hare_f, na.rm = TRUE)

# Get the effect size: 
hare_effsize <- cohen.d(hare_f, hare_m, na.rm = TRUE)
hare_effsize

```

On average, juvenile male hares have larger mean weights than juvenile female hares (`r round(male_mean,2)` $\pm$ `r round(male_sd, 2)` and `r round(female_mean,2)` $\pm$ `r round(female_sd, 2)` g, respectively; mean $\pm$ 1 standard deviation) (Table 1). While the absolute difference in means is `r round(male_mean, 2) - round(female_mean, 2)` g (a `r round((male_mean - female_mean)/((male_mean + female_mean)/2)*100, 2)`% difference), the difference in means is significant (Welch's two-sample t-test: t(`r round(hare_ttest_tidy$parameter,2)`) = `r round(hare_ttest_tidy$statistic,2)`, p < 0.01), and the effect size is small (Cohen's *d* = `r round(hare_effsize$estimate,2)`). There were a small number (*n* = `r hare_weight_stats[3,4]`) of trapped hares without recorded sex. The mean weight for hares labeled "NA" for sex was much lower than for hares recorded as female and male. 


### D. Relationship between juvenile weight & hind foot length

```{r, include = FALSE}
juvenile_foot_weight <- bonanza %>% 
  filter(age == "j") %>% 
  select(weight, hindft)

# Linear regression/model:
hare_lm <- lm(weight ~ hindft, data = juvenile_foot_weight)

# Get tidy versions of the model output to call later on in text
hare_lm_tidy <- tidy(hare_lm)
hare_lm_glance <- glance(hare_lm)

# See model diagnostics for proof of mostly normally distributed and homoscedastic residuals
plot(hare_lm)

# Get Pearson's r correlation
hare_cor <- cor.test(juvenile_foot_weight$weight, juvenile_foot_weight$hindft)

# Tidy version of correlation output:
hare_cor_tidy <- tidy(hare_cor)

summary(hare_lm)
hare_cor

```


```{r}
# Juvenile hare hind foot length versus weight
ggplot(data = juvenile_foot_weight, aes(x = weight, y = hindft)) +
  geom_point(color = "cyan4") +
  labs(x = "Weight (g)", 
       y = "Hind foot length (mm)",
       title = "Juvenile hare hind foot length vs. weight") +
  theme_light() +
  theme(legend.position = "none") +
  geom_smooth(method = "lm",
              color = "black",
              size = 0.5,
              fill = "gray10",
              alpha = 0.5) +
  ggpubr::stat_regline_equation(label.x = 1250, label.y = 93)
```

**Figure 5.** Relationship between flipper length (mm) and body mass (g) for the three penguin species. Points indicate individual penguin measurements (Adélie = gold circles; chinstrap = teal triangles; gentoo = coral diamonds). Linear model summary: $\beta$~1~ = `r round(hare_lm_tidy$estimate[2], 2)` g mm^-1^, p < 0.001, R^2^ = `r round(hare_lm_glance$r.squared,2)`, Pearson's *r* = `r round(hare_cor_tidy$estimate,2)`). Data: Kielland et al. (2017).

In 2 - 3 sentences below your figure, describe the outcome of your linear regression exploration (e.g. slope interpretation, R2 value, and Pearson's r correlation) and touch briefly on which assumption(s) of linear regression may be a concern, and additional thoughts you have about how well (or not well…) the linear model describes the relationship. What worries you? 

The relationship between juvenile hare weight and hind foot length appears relatively linear (Figure 5). Simple linear regression revealed that hare weight significantly predicts hind foot length (p < 0.001, R^2^ = `r round(hare_lm_glance$r.squared,2)`) with an average slope of $\beta$ = `r round(hare_lm_tidy$estimate[2], 2)` g mm^-1^. This means that for each one millimeter increase in hind foot length we expect an average increase in hare weight of `r round(hare_lm_tidy$estimate[2], 2)` g). The Multiple R^2^ value is 0.2988, meaning that 29.9% or ~30% of variance in hind foot length is explained by weight.

From the Pearson's r correlation test, hind foot length and weight were found to be moderately, and significantly, positively correlated (Pearson's *r* = `r round(hare_cor_tidy$estimate,2)`, p < 0.001). Diagnostic plots (not included in this report) reveal relatively normally distributed and homoscedastic residuals. 


## Summary

Exploratory juvenile hares data analysis reveals the following findings:

- Male juvenile hares are the largest of the three species in both mean flipper length and body mass
- Male gentoo penguins longer mean flipper length than female gentoos (the difference is significant, and the effect size is large)
- Flipper length appears linearly related with body mass across all three penguin species; the relationship (by simple linear regression) is significant, with a slope (expected average increase in body mass with a 1 mm increase in flipper length) of 49.69 g mm-1.

Next steps:

- Explore the impacts of sex and site when comparing the relationship between weight and hind foot length.


## Citations

[Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative.](https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14)

